A sheaf--function dictionary for algebraic tori over local fields

### 2013-09-19 shanghai.tex
Modifications to the abstract and introduction; please look at them. Other small changes ....
Fixed ell at the beginning of the paper (we seem to have forgotten do
do that). Typo in Finite QC Sheaves. All other changes are documented
in the file (comments).

\begin{abstract}
  We introduce and study the category of quasicharacter sheaves on smooth,
  commutative group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of quasicharacter sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between quasicharacters of $T(K)$ and isomorphism
  classes of quasicharacter sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-archimedean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{tate:67a}*{\S 2.3}
%\end{quotation}
%\noindent
In this paper we present a dictionary between quasicharacters of algebraic tori over local fields and quasicharacter sheaves.
What is a quasicharacter sheaf?
Let $G$ be a smooth, commutative group scheme locally of finite
type over a finite field $\Fq$; $G$ need not be reductive, connected, or of
finite type over $\Fq$. Fix a prime $\ell$ invertible in $\Fq$
and write $m : \GxG \to G$ for the multiplication morphism,
$\bG$ for the base change of $G$ to $\bFq$ and
$\Frob{G} : \bG \to \bG$ for the Frobenius automorphism.
A quasicharacter sheaf on $G$ is a triple $\qcs{L}\ceq
(\gqcs{L},\mu,\phi)$ where $\gqcs{L}$ is an $\ell$-adic local system on $\bG$ of rank-1 and $\mu: m^*
\gqcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gqcs{L} \to \gqcs{L}$ are isomorphisms;
$\mu$ and $\phi$ are required to satisfy natural conditions explained below.

In Part 1 of this paper we study the category $\QC(G)$ of quasicharacter sheaves.
Since it is a rigid symmetric monoidal category
the set of isomorphism classes $\QCiso{G}$ forms a group.
Motivated by the sheaf--function dictionary for commutative, connected
algebraic groups \cite{deligne:SGA4.5}*{sommes trig.}, 
we explore the trace of Frobenius morphism
\[
\QCiso{G} \to \Hom(G(\Fq), \EEx)
\]
and show that it is an isomorphism in Theorem~\ref{thm:snake}, assuming only
that the geometric component group of $G$ is finitely generated.

In Part 2 we use Theorem~\ref{thm:snake} to make a sheaf--function dictionary
for algebraic tori over local fields.  Let $K$ be a non-archimedian local field
with finite residue field $\Fq$; we place no restriction on the characteristic of $K$.
Write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$.
Let $T$ be a torus defined over $K$.  The N\'eron model $\TT$ of $T$
is a smooth group scheme over $\OK$ with generic fiber $T$ and a canonical isomorphism
$\TT(\OK) \cong T(K)$.  For each positive integer $n$, let $\GN{T}_n$ be the
Greenberg transform of $\TT \times_{\Spec{\OK}} \Spec{\OK/\pK^n}$; the commutative
group scheme $\GN{T}_n$ plays the r\^{o}le of $G$.  Theorem~\ref{thm:application}
allows us to translate characters $T(K)$ of depth less than $n$ into quasicharacter sheaves on $\GN{T}_n$:
\[
\QCiso{\GN{T}_n} 
\iso
\Hom_{< n}(T(K),\EEx).
\]  
We refer to finite-depth characters of $T(K)$ as quasicharacters and
describe a pro-group scheme $\GN{T}$ simultaneously geometrizing all such quasicharacters.

We have not yet fully explored the applications of such a geometrization, but want to highlight
some of the potential benefits of Theorem~\ref{thm:application}.  Characters on $T$ have limited
functoriality: they can be pulled back along maps to $T$, but it is less clear what to do with them
in the presence of maps from $T$ to other reductive groups.  Sheaves, on the other hand, come
equipped with a host of functors, including pushforward.

DeBacker and Reeder have recently applied Deligne-Lusztig representations to part of the
local Langlands correspondence. % \cite{reeder-debacker:09a}.  
After developing Deligne-Lusztig representations, 
the next step Lusztig took in studying representations
of finite groups of Lie type was to introduce character sheaves for reductive groups over $\Fq$.
We hope that our quasicharacter sheaves can provide a stepping stone to a similar theory for
representations of $p$-adic groups.

@@

For any group $A$, we will denote by $A^*$ the dual group $\Hom_\text{grp}(A, \EEx)$.

@@

To define the category of finite quasicharacter sheaves we must work a bit harder.
Consider the following category $C(G)$.
Let us say that $f : H\to G$ is a \emph{discrete isogeny} if it is a finite, surjective
\'etale morphism and a morphism of commutative group $\Fq$-schemes, for which
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $f^{-1}({\bar e})$ is trivial.
An object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is a discrete isogeny 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G
\end{tikzcd}
\]
We say that $(\alpha,T)$ in $C(G)$ is a \emph{weak isomorphism} if $\alpha$ is surjective
and $T$ is an isomorphism of vector spaces.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.

@@

\begin{proof}
The key point is that pull-back along $g$ takes isogenies to isogenies,
and maps over $G$ to maps over $G'$, as pictured below.
\[
\begin{tikzcd}[row sep=10,column sep=10]
\ & V \arrow{dd}{T} && \Aut(V) && \arrow[swap]{ll}{\psi} \ker f \arrow[hook]{rr} && H \arrow{rr}{f} && \arrow[equal]{dd} G\\
 && && \arrow{ul} \ker f\times_G G' \arrow{ur} \arrow[hook]{rr} && H\times_G G' \arrow{ur} \arrow{rr} && G' \arrow{ur}{g} & \\
& V' && \Aut(V) && \arrow[swap,pos=0.7]{ll}{\psi'} \arrow[swap]{uu}{} \ker f'  \arrow[hook]{rr}
&& \arrow[swap]{uu} H' \arrow[pos=0.8]{rr}{f'} && G\\
 && && \arrow{ul} \ker f'\times_G G' \arrow{uu} \arrow{ur} \arrow[hook]{rr}
 && \arrow{uu} H'\times_G G' \arrow{ur} \arrow{rr} && \arrow[equals]{uu} G' \arrow[swap]{ur}{g} & 
\end{tikzcd}
\]
In this form we could treat $g^*$ as a functor $G(G')\to C(G)$,
where it not for the fact that pull-back is defined only up to isomorphism.
In the localized categories, this ambiguity disappears. 
\end{proof}

@@

  Since the action of $\Gal(\bFq/\Fq)$ is continuous, there is an $m$
  so that $\Frob{\Fq}^m$ acts trivially on $X$; write $\Fqm$ for the degree
  $m$ extension of $\Fq$ and $\Gamma_m$ for $\Gal(\Fqm/\Fq)$.  Since
  $T^m-1$ factors as a squarefree product of cyclotomic polynomials,
  the Chinese remainder theorem expresses $\ZZ[\Gamma_m]$ as a product
  of $\ZZ[\zeta_d]$ for $d$ dividing $m$.  Each $\ZZ[\zeta_d]$ is the
  maximal order in a number field and thus $\ZZ[\Gamma_m]$ is a
  product of Dedekind domains.  Therefore $X$ decomposes into a direct
  sum of cyclic $\ZZ[\Gamma_m]$-modules \cite{brandal:CommRingsFGModDecompose}, \David{more specific citation}
  each of the \David{Add a reference to this section}
  form $\ZZ[\Gamma_m] / I$ for some ideal $I \subset \ZZ[\Gamma_m]$.
  The ideal $I$ corresponds to a direct product of ideals $I_d$ in the
  $\ZZ[\zeta_d]$, and within each $\ZZ[\zeta_d]$ the Chinese remainder
  theorem allows a further decomposition of $\ZZ[\zeta_d] / I_d$ as a
  direct sum of $\ZZ[\zeta_d]$-modules of the form $\ZZ[\zeta_d]/P^s$.
  So if we can construct, for every cyclotomic ring $\ZZ[\zeta_d]$,
  prime ideal $P \subset \ZZ[\zeta_d]$ and positive integer $s$, a
  $\psi \colon \ZZ[\zeta_d]/P^s \rightarrow \EEx$ with
  \begin{equation} \label{eq:psi-condition}
    \tilde{f}(x) = \frac{\psi(\zeta_d \cdot x)}{\psi(x)},
  \end{equation}
  then the existence of the original $\psi$ follows because $X$ is a
  direct sum of such $\ZZ[\zeta_d]/P^s$.

@@

  The relations among the generators $\zeta^i$ are additively
  generated by $\zeta^ch(\zeta)$ as $c$ ranges from $0$ to
  $\phi(d)-r-1$.  If we set
  $$A_c = \sum_{i=0}^r a_i \sum_{j=0}^{i+c-1} \zeta^j$$
  then
  $$\psi(\zeta^ch(\zeta)) = \alpha^{h(1)} \tilde{f}(A_c).$$
  Now consider $(\zeta - 1)A_c$: the sums telescope and we get that
  $(\zeta-1)A_c = \zeta^ch(\zeta) - h(1).$ There are now two
  possibilities:
  \begin{enumerate}
  \item $d$ is a power of $p$; write $s = \phi(d)m + n$.  We may take
    $h(T)$ to be $p^m(T-1)^n$, in which case $\zeta^ch(\zeta) - h(1) \in P^s$,
    so $A_c$ is fixed under multiplication by $\zeta$ in
    $\ZZ[\zeta]/P^s$ and thus $\tilde{f}(A_c) = 1$.  In this case
    $\alpha^{h(1)} = 1$ regardless of $\alpha$.
  \item $d$ is not a power of $p$.  If two distinct primes divide $d$
    then $\zeta - 1$ is a unit; otherwise $\zeta - 1$ is a unit modulo
    $P^s$.  Let $\omega \in \ZZ[\zeta]$ satisfy
    $\omega(\zeta - 1) \equiv 1 \pmod{P^s}$ and set
    \begin{equation} \label{alpha-def}
      \alpha = \tilde{f}(\omega).
    \end{equation}
    Then $A_c + h(1)\omega$ is fixed under multiplication by $\zeta$
    and thus $\tilde{f}(A_c) = \tilde{f}(-h(1)\omega)$.
  \end{enumerate}
  In either case, $\psi(\zeta^ch(\zeta)) = 1$ for all $c$.  The only
  other relations imposed on the $\zeta^i$ are that they are all
  killed under multiplication by $p^s$.  Since $\tilde{f}$ is also a
  homomorphism $\ZZ[\zeta]/P^s \rightarrow \EEx$ the values of
  $\tilde{f}$ and thus of $\psi$ will be $p^s$-roots of unity.
  Therefore $\psi$ is well defined.

  We now check Equation~\eqref{eq:psi-condition} for
  $x = \zeta^{\phi(d)-1}$.  Write
  $\Phi_d(T) = T^{\phi(d)} - b_{\phi(d)-1}T^{\phi(d)-1} - \cdots - b_0$
  and set
  $$B = \sum_{i=0}^{\phi(d)-1} -\zeta^i + b_i \sum_{j=0}^{i-1} \zeta^j.$$
  We have

  \begin{align*}
    \frac{\psi(\zeta^{\phi(d)})}{\psi(\zeta^{\phi(d)-1})\tilde{f}(\zeta^{\phi(d)-1})}
    &= \frac{\psi(b_{\phi(d)-1}\zeta^{\phi(d)-1} + \cdots + b_0)}{\psi(\zeta^{\phi(d)-1})\tilde{f}(\zeta^{\phi(d)-1})} \\
    &= \alpha^{-\Phi_d(1)} \tilde{f}(B)
  \end{align*}
  As before we see that $(\zeta - 1)B = \Phi_d(1) - \Phi_d(\zeta) = \Phi_d(1)$.
  We consider the same cases as before:

@@

\begin{corollary}
 If $G$ is a commutative, \'etale group scheme over $\Fq$ and $G(\bFq)$ is finitely
 generated then
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{corollary}
\begin{proof}
This follows from Proposition~\ref{prop:etale} by setting $X=G(\bFq)$ in Proposition~\ref{prop:X}.
\end{proof}

@@

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0} in the category of $\QC$-schemes
  defining the component group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker \trFrob{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker \trFrob{G^0} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\trFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\trFrob{G}} & \QCiso{G^0} \rar \dar{\trFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker \trFrob{\pi_0(G)} & \coker \trFrob{G} &  \coker \trFrob{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  Since $k$ is perfect
  \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}} implies that
  \David{Need another reference that an algebraic group over an algebraically closed field is reduced}
  $G$ is reduced.  Since $G^0$ is a finite type
  \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}, it is thus a connected algebraic group and 
  Proposition~\ref{prop:connected} implies that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

>>>>

\begin{abstract}
  We introduce and study the category of quasicharacter sheaves on smooth,
  commutative group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of quasicharacter sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. Then, for any non-archimedean local field $K$
  and any algebraic torus $T$ over $K$, we use this result to produce a functorial
  isomorphism between quasicharacters of $T(K)$ and isomorphism
  classes of quasicharacter sheaves on the Greenberg transform of the Néron
  model of $T$.
  \end{abstract}

@@

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{tate:67a}*{\S 2.3}
%\end{quotation}
%\noindent
In this paper we find a dictionary between quasicharacters of algebraic tori over local fields and quasicharacter sheaves.

What is a quasicharacter sheaf?
Let $G$ be a smooth, commutative group scheme locally of finite
type over a finite field $\Fq$; $G$ need not be reductive, connected, or of
finite type over $\Fq$. Fix a prime $\ell$ invertible in $\Fq$
and write $m : \GxG \to G$ for the multiplication morphism,
$\bG$ for the base change of $G$ to $\bFq$ and
$\Frob{G} : \bG \to \bG$ for the Frobenius automorphism.
A quasicharacter sheaf on $G$ is a triple $\qcs{L}\ceq
(\gqcs{L},\mu,\phi)$ where $\gqcs{L}$ is an $\ell$-adic local system on $\bG$ of rank-1 and $\mu: m^*
\gqcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gqcs{L} \to \gqcs{L}$ are isomorphisms;
$\mu$ and $\phi$ are required to satisfy natural conditions explained below.

In Part 1 of this paper we study the category $\QC(G)$ of quasicharacter sheaves on $G$.
Since $\QC(G)$ is a rigid symmetric monoidal category
the set of isomorphism classes $\QCiso{G}$ forms a group.
Motivated by the sheaf--function dictionary for commutative, connected
algebraic groups \cite{deligne:SGA4.5}*{sommes trig.}, 
we explore the trace of Frobenius morphism
\[
\QCiso{G} \to \Hom(G(\Fq), \EEx)
\]
and show in Theorem~\ref{thm:snake} that it is an isomorphism, 
assuming that the geometric component group of $G$ is finitely generated.

In Part 2 we use Theorem~\ref{thm:snake} to make a sheaf--function dictionary
for algebraic tori over local fields.  Let $K$ be a non-archimedian local field
with finite residue field $\Fq$; we place no restriction on the characteristic of $K$.
Write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$.
Let $T$ be a torus defined over $K$.  The N\'eron model $\TT$ of $T$
is a smooth group scheme over $\OK$ with generic fibre $T$ and a canonical isomorphism
$\TT(\OK) \cong T(K)$.  For each positive integer $n$, let $\GN{T}_n$ be the
Greenberg transform of $\TT \times_{\Spec{\OK}} \Spec{\OK/\pK^n}$; the commutative
group scheme $\GN{T}_n$ plays the r\^{o}le of $G$.  Theorem~\ref{thm:application}
allows us to translate characters $T(K)$ of depth less than $n$ into quasicharacter sheaves on $\GN{T}_n$:
\[
\QCiso{\GN{T}_n} 
\iso
\Hom_{< n}(T(K),\EEx).
\]  
We also describe a pro-group scheme $\GN{T}$ simultaneously geometrizing all quasicharacters ({\it i.e.}, finite-depth characters) of $T(K)$.

While we have not yet fully explored the applications of such a geometrization, we want to highlight
some of the potential benefits of Theorem~\ref{thm:application} here. Characters on $T$ have limited
functoriality: they can be pulled back along maps to $T$, but it is less clear what to do with them
in the presence of maps from $T$ to other groups.  Sheaves, on the other hand, come
equipped with a host of functors, including pushforward. 
We hope to use this to our advantage as we work toward an adaptation of Lusztig's theory of characters sheaves to the study of characters of admissible representations of reductive $p$-adic groups that extends beyond the case of $p$-adic tori.

DeBacker and Reeder have recently applied Deligne-Lusztig representations to part of the
local Langlands correspondence. % \cite{reeder-debacker:09a}.  
After developing Deligne-Lusztig representations, 
the next step Lusztig took in studying representations
of finite groups of Lie type was to introduce character sheaves for reductive groups over $\Fq$.
As suggested above above, we hope that our quasicharacter sheaves 
can provide a stepping stone to a similar theory for
representations of $p$-adic groups.

@@

We fix a prime $\ell$, invertible in $\Fq$.
For any group $A$, we will denote by $A^*$ the dual group $\Hom_\text{grp}(A, \EEx)$.

@@

To define the category of finite quasicharacter sheaves we must work a bit harder.
Consider the following category $C(G)$.
Let us say that $f : H\to G$ is a \emph{discrete isogeny} if it is a finite, surjective
\'etale morphism and a morphism of commutative group $\Fq$-schemes for which
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $f^{-1}({\bar e})$ is trivial.
An object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is a discrete isogeny 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G
\end{tikzcd}
\]
We say that $(\alpha,T)$ in $C(G)$ is a \emph{weak isomorphism} if $\alpha$ is surjective
and $T$ is an isomorphism of vector spaces.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.

@@

\begin{proof}
The key point is that pull-back along $g$ takes isogenies to isogenies,
and maps over $G$ to maps over $G'$, as pictured below.
\[
\begin{tikzcd}[row sep=10,column sep=10]
\ & V \arrow{dd}{T} && \Aut(V) && \arrow[swap]{ll}{\psi} \ker f \arrow[hook]{rr} && H \arrow{rr}{f} && \arrow[equal]{dd} G\\
 && && \arrow{ul} \ker f\times_G G' \arrow{ur} \arrow[hook]{rr} && H\times_G G' \arrow{ur} \arrow{rr} && G' \arrow{ur}{g} & \\
& V' && \Aut(V) && \arrow[swap,pos=0.7]{ll}{\psi'} \arrow[swap]{uu}{} \ker f'  \arrow[hook]{rr}
&& \arrow[swap]{uu} H' \arrow[pos=0.8]{rr}{f'} && G\\
 && && \arrow{ul} \ker f'\times_G G' \arrow{uu} \arrow{ur} \arrow[hook]{rr}
 && \arrow{uu} H'\times_G G' \arrow{ur} \arrow{rr} && \arrow[equals]{uu} G' \arrow[swap]{ur}{g} & 
\end{tikzcd}
\]
In this form we could treat $g^*$ as a functor $C(G')\to C(G)$,
where it not for the fact that pull-back is defined only up to isomorphism.
In the localized categories, this ambiguity disappears. 
\end{proof}

@@

  Since the action of $\Gal(\bFq/\Fq)$ is continuous, there is an $m$
  so that $\Frob{\Fq}^m$ acts trivially on $X$; write $\Fqm$ for the degree
  $m$ extension of $\Fq$ in $\bFq$ and $\Gamma_m$ for $\Gal(\Fqm/\Fq)$.  Since
  $T^m-1$ factors as a squarefree product of cyclotomic polynomials,
  the Chinese remainder theorem expresses $\ZZ[\Gamma_m]$ as a product
  of $\ZZ[\zeta_d]$ for $d$ dividing $m$.  Each $\ZZ[\zeta_d]$ is the
  maximal order in a number field and thus $\ZZ[\Gamma_m]$ is a
  product of Dedekind domains.  Therefore $X$ decomposes into a direct
  sum of cyclic $\ZZ[\Gamma_m]$-modules \cite{brandal:CommRingsFGModDecompose}, \David{more specific citation} \Clifton{I'm tracking this down.}
  each of the \David{Add a reference to this section}
  form $\ZZ[\Gamma_m] / I$ for some ideal $I \subset \ZZ[\Gamma_m]$.
  The ideal $I$ corresponds to a direct product of ideals $I_d$ in the
  $\ZZ[\zeta_d]$, and within each $\ZZ[\zeta_d]$ the Chinese remainder
  theorem allows a further decomposition of $\ZZ[\zeta_d] / I_d$ as a
  direct sum of $\ZZ[\zeta_d]$-modules of the form $\ZZ[\zeta_d]/P^s$.
  So if we can construct, for every cyclotomic ring $\ZZ[\zeta_d]$,
  prime ideal $P \subset \ZZ[\zeta_d]$ and positive integer $s$, a
  $\psi \colon \ZZ[\zeta_d]/P^s \rightarrow \EEx$ with
  \begin{equation} \label{eq:psi-condition}
    \tilde{f}(x) = \frac{\psi(\zeta_d \cdot x)}{\psi(x)},
  \end{equation}
  then the existence of the original $\psi$ follows because $X$ is a
  direct sum of such $\ZZ[\zeta_d]/P^s$.

@@

  The relations among the generators $\zeta^i$ are additively
  generated by $\zeta^ch(\zeta)$ as $c$ ranges from $0$ to
  $\phi(d)-r-1$.  If we set
  $$A_c = \sum_{i=0}^r a_i \sum_{j=0}^{i+c-1} \zeta^j$$
  then
  $$\psi(\zeta^ch(\zeta)) = \alpha^{h(1)} \tilde{f}(A_c).$$
  Now consider $(\zeta - 1)A_c$: the sums telescope and we get that
  $(\zeta-1)A_c = \zeta^ch(\zeta) - h(1).$ There are now two
  possibilities:
  \begin{enumerate}
  \item $d$ is a power of $p$; write $s = \phi(d)m + n$.  We may take
    $h(T)$ to be $p^m(T-1)^n$, in which case $\zeta^ch(\zeta) - h(1) \in P^s$,
    so $A_c$ is fixed under multiplication by $\zeta$ in
    $\ZZ[\zeta]/P^s$ and thus $\tilde{f}(A_c) = 1$.  In this case
    $\alpha^{h(1)} = 1$ regardless of $\alpha$.
  \item $d$ is not a power of $p$.  If two distinct primes divide $d$
    then $\zeta - 1$ is a unit; otherwise $\zeta - 1$ is a unit modulo
    $P^s$.  Let $\omega \in \ZZ[\zeta]$ satisfy
    $\omega(\zeta - 1) \equiv 1 \pmod{P^s}$ and set
    %\begin{equation} \label{alpha-def}
     $ \alpha = \tilde{f}(\omega)$.
    %\end{equation}
    Then $A_c + h(1)\omega$ is fixed under multiplication by $\zeta$
    and thus $\tilde{f}(A_c) = \tilde{f}(-h(1)\omega)$.
  \end{enumerate}
  In either case, $\psi(\zeta^ch(\zeta)) = 1$ for all $c$.  The only
  other relations imposed on the $\zeta^i$ are that they are all
  killed under multiplication by $p^s$.  Since $\tilde{f}$ is also a
  homomorphism $\ZZ[\zeta]/P^s \rightarrow \EEx$ the values of
  $\tilde{f}$ and thus of $\psi$ will be $p^s$-roots of unity.
  Therefore $\psi$ is well defined.

  We now check Equation~\eqref{eq:psi-condition} for
  $x = \zeta^{\phi(d)-1}$.  Write
  \[\Phi_d(T) = T^{\phi(d)} - b_{\phi(d)-1}T^{\phi(d)-1} - \cdots - b_0\]
  and set
  $$B = b_i \sum_{j=0}^{i-1} \zeta^j -  \sum_{i=0}^{\phi(d)-1} \zeta^i.$$
  We have

  \begin{align*}
    \frac{\psi(\zeta^{\phi(d)})}{\psi(\zeta^{\phi(d)-1})\tilde{f}(\zeta^{\phi(d)-1})}
    &= \frac{\psi(b_{\phi(d)-1}\zeta^{\phi(d)-1} + \cdots + b_0)}{\psi(\zeta^{\phi(d)-1})\tilde{f}(\zeta^{\phi(d)-1})} 
    = \alpha^{-\Phi_d(1)} \tilde{f}(B).
  \end{align*}
  As before we see that $(\zeta - 1)B = \Phi_d(1) - \Phi_d(\zeta) = \Phi_d(1)$.
  We consider the same cases as before:

@@

\begin{corollary}
 If $G$ is a commutative, \'etale group scheme over $\Fq$ and $G(\bFq)$ is finitely
 generated then
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{corollary}
\begin{proof}
This follows from Proposition~\ref{prop:etale} by setting $X=G(\bFq)$ in Proposition~\ref{prop:X}.
Recall that $G \mapsto G(\bFq)$ defines an equivalence of categories from commutative \'etale group schemes over $\Fq$ to commutative groups equipped with \emph{continuous} actions by $\Gal(\bFq/\Fq)$; see, for example, \cite{vdGeer-Moonen:AbelianVarieties}*{Prop 3.26}.
\end{proof}

@@

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0} in the category of $\QC$-schemes
  defining the component group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker \trFrob{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker \trFrob{G^0} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\trFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\trFrob{G}} & \QCiso{G^0} \rar \dar{\trFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker \trFrob{\pi_0(G)} & \coker \trFrob{G} &  \coker \trFrob{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.

   We claim that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$. 
   This follows from Proposition~\ref{prop:connected} 
   once we confirm that $G^0$ is an algebraic group (it is clearly connected).
   To see this, note that $G^0$ is a smooth group scheme of finite type over $\Fq$ by \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
   Since $k$ is perfect, $G^0$ is also reduced, by
  \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}}.
  \David{Need another reference that an algebraic group over an algebraically closed field is reduced}
  \Clifton{I'm working on this too.}
  Thus, $G^0$ is indeed an algebraic group.

  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.
